Cooling Requirements for the Vertical Shear Instability in Protoplanetary Disks

### 2015-05-08 appendix.tex
re-shuffled 6.4, updated fig 2, poster files update

Consider the simplified dispersion relations for marginal stability,
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. We write 
$X= \hat{\omega}\khat$, $\theta = (hq\khat)^2$ and  
treat $M$ as a continuous variable. We find from
Eq. \ref{relax_cond_simp1} that  
%\begin{align}
%  &x^4  = x^2 + M(M+1)(1-\theta), \label{simp1}\\
%  &x^2  = \frac{\gamma-1}{h q}\beta_c (1+2M)^2x - 2M(M+1)\label{simp2},
%\end{align}
%where and
\begin{align}
  2 X \frac{d X}{dM} = \frac{X^2(1-\theta)(2M+1)}{X^2 +
  2M(M+1)(1-\theta)}, 
\end{align}
and from Eq. \ref{relax_cond_simp2} that
\begin{align}
  2X\frac{dX}{dM} = \left[X^2 +
  2M(M+1)\right]\left(\frac{d\ln{\beta_c}}{dM} + \frac{4}{2M+1} +
  \frac{d\ln{X}}{dM}\right) - 2(2M+1). 
\end{align}
We eliminate $dX/dM$, making use of Eq. \ref{relax_cond_simp1} in the process, to obtain 
\begin{align}
  (2M+1)\frac{d\ln{\beta_c}}{dM} %= \frac{\mathcal{N}}{\mathcal{D}}, 
  = -\frac{(1-\theta)M(M+1)\left[ 2(2M+1)^2 + 12X^2\right]
  + (3+\theta)X^2}{2 \left[X^2 +
  2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]}.  
\end{align}
%where the denominator $\mathcal{D}$ is
%\begin{align}
%\mathcal{D} = 2 \left[X^2 +
 % 2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]>0, 
%\end{align}
%and the numerator $\mathcal{N}$ is 
%\begin{align}
 % \mathcal{N} = -(1-\theta)\left[ 2M(M+1)(2M+1)^2 + 12X^2M(M+1)\right]
 % - (3+\theta)X^2 < 0.
%\end{align}
Hence,
\begin{align}\label{dbeta_dm}
  \frac{d\beta_c}{dM}<0, 
\end{align}
for all $M\geq 0$ if $\theta\leq 1$, which imply 
$\mathrm{max}(\beta_c)$ occurs at $M=0$. This result may also be
obtained by explicitly solving 
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$
treated as a small parameter. For fixed $\khat$ the 
condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
All all such modes are stabilized if
$\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$.

This result highlights the importance of the fundamental mode  
--- it is the most difficult mode to stabilize with finite
cooling. Furthermore, for $M=0$ we may obtain the expression  
for $\beta_\mathrm{crit}$ from the dispersion relations 
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} without  
assuming it is much less than unity from the onset or place
restrictions on $\theta$.

>>>>

Consider the simplified dispersion relations for marginal stability,
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. We write 
$X= \hat{\omega}\khat$, $\theta = (hq\khat)^2$ and  
treat $M$ as a continuous variable. We find from
Eq. \ref{relax_cond_simp1} that  
%\begin{align}
%  &x^4  = x^2 + M(M+1)(1-\theta), \label{simp1}\\
%  &x^2  = \frac{\gamma-1}{h q}\beta_c (1+2M)^2x - 2M(M+1)\label{simp2},
%\end{align}
%where and
\begin{align}
  2 X \frac{d X}{dM} = \frac{X^2(1-\theta)(2M+1)}{X^2 +
  2M(M+1)(1-\theta)}, 
\end{align}
and from Eq. \ref{relax_cond_simp2} that
\begin{align}
  2X\frac{dX}{dM} = \left[X^2 +
  2M(M+1)\right]\left(\frac{d\ln{\beta_c}}{dM} + \frac{4}{2M+1} +
  \frac{d\ln{X}}{dM}\right) - 2(2M+1). 
\end{align}
We eliminate $dX/dM$, making use of Eq. \ref{relax_cond_simp1} in the process, to obtain 
\begin{align}
  (2M+1)\frac{d\ln{\beta_c}}{dM} %= \frac{\mathcal{N}}{\mathcal{D}}, 
  = -\frac{(1-\theta)M(M+1)\left[ 2(2M+1)^2 + 12X^2\right]
  + (3+\theta)X^2}{2 \left[X^2 +
  2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]}.  
\end{align}
%where the denominator $\mathcal{D}$ is
%\begin{align}
%\mathcal{D} = 2 \left[X^2 +
 % 2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]>0, 
%\end{align}
%and the numerator $\mathcal{N}$ is 
%\begin{align}
 % \mathcal{N} = -(1-\theta)\left[ 2M(M+1)(2M+1)^2 + 12X^2M(M+1)\right]
 % - (3+\theta)X^2 < 0.
%\end{align}
Hence,
\begin{align}\label{dbeta_dm}
  \frac{d\beta_c}{dM}<0, 
\end{align}
for all $M\geq 0$ if $\theta\leq 1$, which imply 
$\mathrm{max}(\beta_c)$ occurs at $M=0$. This result may also be
obtained by explicitly solving 
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$ as a small parameter. 
For fixed $\khat$ the condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
All such modes are stabilized if
$\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$.

This result highlights the importance of the fundamental mode  
--- it is the most difficult mode to stabilize with finite
cooling. Furthermore, for $M=0$ we may obtain the expression  
for $\beta_\mathrm{crit}$ from the dispersion relations 
Eq. \ref{relax_cond1}---\ref{relax_cond2} without  
assuming it is much less than unity from the onset or place
restrictions on $\theta$.

### 2015-05-08 summary.tex
minor edits 2,3,4
minor to 3.4

\section{Caveats and Neglected Effects} 
%We highlight below some outstanding issues that should be addressed in
%future works.

@@

\section{Summary and discussion}\label{summary}
In this paper, we have studied the vertical shear instability (VSI) with a focus the 
role of radiative cooling.  In turn we assess the viability of the VSI 
as an angular momentum transport mechanism in protoplanetary disks (PPDs).

>>>>

\section{Caveats and Neglected Effects}\label{caveats} 
%We highlight below some outstanding issues that should be addressed in
%future works.

@@

\section{Summary and discussion}\label{summary}
In this paper we study the vertical shear instability (VSI) with a focus the 
role of radiative cooling.  In turn we assess the viability of the VSI 
as an angular momentum transport mechanism in protoplanetary disks (PPDs).
